---
title: "Project_Summary"
author: "Diego Uchendu"
date: "25/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(kableExtra)
library(knitr)
library(gtools)
```
```{r}
library(ROI)
library(ROI.plugin.alabama)
```

# Data

```{r}
library(readxl)
```


```{r}
product_data <- read_excel("Data_ordering.xlsx",sheet = "product data")
box_data <- read_excel("Data_ordering.xlsx",sheet = "box data")
```


```{r}
#rack data
Total_racks=8
levels_per_rack=4
rack_length= 6000
rack_width=1750
rack_height=300
```




```{r}
Racks<- data.frame(Total_racks,levels_per_rack,rack_length,rack_width,rack_height)
colnames(Racks)<- c("Total_racks","levels_per_rack","length(mm)","rack_width(mm)","rack_height(mm)")
rownames(Racks)<-"Rack Information"

Racks
#kable(Racks,digits = c(1,1,1,1,1), caption= "Storage Information", format= "pandoc")
```

## Box data

```{r}
box_data
```


## Product data 

```{r}
#product_data
product_data<-cbind(1:62,product_data)
colnames(product_data)<- c("N0", "material ID","demand per day","box ID","pieces/box","price")

#kable(product_data,format = "pandoc")
```




```{r}

product_data$demand_per_year= ceiling((product_data$`demand per day` *262)/ product_data$`pieces/box`) 

product_data$box_cost= product_data$`pieces/box`*product_data$price
```




```{r}
product_data <- merge(product_data, box_data[,c("box ID", "ordering cost (€)")], by = "box ID")

colnames(product_data)[9] <- "ordering_cost"


#kable(product_data,digits = c(1,1,1,1,1,1,1,1), caption= "Merged data with yearly demand", format= "pandoc")

#product_data<- product_data[,c("N0", "material ID","demand per day","box ID","pieces/box","price")]
#product_data
```


```{r}
product_data<- product_data[order(product_data$N0),]

row.names(product_data)<-1:62
```


```{r}
#kable(product_data,format = "pandoc")


product_data
```



## EOQ



```{r}
#dei= demand_per_year, cori= ordering cost for box i
#cord= ordering cost, pri= box cost, h= interest rate

# directly vectorized
so_eoq_fun <- Vectorize(function(dei,cori,cord,pri,h){
  eoq<- sqrt((2*dei*(cori+cord))/(pri*h))
  return(eoq)
})

# maximum EOQ provided there is no coordination of ordering cycles (i.e., cord are not shared among parallely ordered items)
vec_so_eoq_fun.max <- so_eoq_fun(dei = product_data$demand_per_year,cori=product_data$ordering_cost,cord=1500,pri=product_data$box_cost,h=0.10)
# minimum EOQ disregarding common ordering cost cord
vec_so_eoq_fun.min <- so_eoq_fun(dei = product_data$demand_per_year,cori=product_data$ordering_cost,cord=0,pri=product_data$box_cost,h=0.10)

product_data$eoq.min <- round(vec_so_eoq_fun.min)
product_data$eoq.max <- round(vec_so_eoq_fun.max)
#product_data

#kable(product_data,digits = c(1,1,1,1,1,1,1,1), caption= "Both EOQs added to the table", format= "pandoc")

product_data
```


## Lane Occupation


```{r}
# constraints #####################################
# this can be formulated more elegantly, but it works and that suffices
b_sorting <- double(length(product_data$`box ID`)) 
b_not_sorting <-double(length(product_data$`box ID`))

for(j in 1: length(box_data$`box ID`)){
  for (k in 1:length(product_data$`box ID`)) {
    if(box_data$`box ID`[j]==product_data$`box ID`[k]){
      
      if(box_data$sorting[j]=="width"){
        b_sorting[k] <- box_data$width[j]
        b_not_sorting[k]<- box_data$length[j]
        
      }else{
        b_sorting[k] <- box_data$length[j]
        b_not_sorting[k]<- box_data$width[j]
      }
    }
  }
  
}

product_data$b_sorting <- b_sorting 
product_data$b_not_sorting <- b_not_sorting
#product_data

#kable(product_data,digits = c(1,1,1,1,1,1,1,1), caption= "Both EOQs added to the table", format= "pandoc")

product_data

```
## Lane EOQ max and min

```{r}
lane <- ceiling(product_data$eoq.max * (product_data$b_not_sorting/rack_length))
laneso<- data.frame(rbind(lane))
rownames(laneso)<- "lanes"
colnames(laneso)<- 1:62
#kable(laneso,format = "pandoc",caption = "number of lanes per item")

rack_total_width <-51200  #rack_width * 4 * 8
#rack_total_width #56000

lane.min <- ceiling(product_data$eoq.min * product_data$b_not_sorting / rack_length)
#lane.min
lane.max <- ceiling(product_data$eoq.max * product_data$b_not_sorting / rack_length)

#lane.max

# total rack width 

rack_total_width <- 51200 # rack_width * 4 * 8

lane.df<- data.frame(rbind(lane.min,lane.max))
rownames(lane.df)<-c("Min lane","Max lane")
colnames(lane.df)<-1:62
#kable(lane.df,caption = "Number of lanes that minimizes cost",format = "pandoc")

lane.df


```

## Ignor displaying

```{r}
# overflow in mm
overflow.mm.lane.min<- sum(lane.min*product_data$b_sorting) - rack_total_width
overflow.mm.lane.max<-sum(lane.max*product_data$b_sorting) - rack_total_width
# relative overflow in %
opt_cap_in_mm <- sum(lane.min*product_data$b_sorting)

overflow.per.lane.min<-(sum(lane.min*product_data$b_sorting) - rack_total_width)/rack_total_width*100
overflow.per.lane.max<- (sum(lane.max*product_data$b_sorting) - rack_total_width)/rack_total_width*100
n0_same_cap_needed<- (sum(lane.max*product_data$b_sorting) - rack_total_width)/rack_total_width
overflow<- data.frame(overflow.mm.lane.min,overflow.mm.lane.max,overflow.per.lane.min,overflow.per.lane.max,n0_same_cap_needed)

colnames(overflow)<-c("mm_lane_min","mm_lane_max","%_lane_min","%_lane.max","n0_same_cap_needed")
rownames(overflow)<- "Overflow"

#kable(overflow,caption = "Rack Capacity overflow in mm and percentage",format = "pandoc")

#product_data
```


```{r}
#get the width of the lanes in mm
rack_width_occupied=sum(lane.min*product_data$b_sorting)
kable(paste(rack_width_occupied,"mm"," is the total expected to be occupied"))

if(rack_width_occupied <= rack_total_width){
  print(paste0("Capacity constraint fulfilled: ", rack_width_occupied, "<=",rack_total_width))
 
}else {
  violated <- rack_width_occupied - rack_total_width

## kable((paste0("Capacity constraint violated by: ",violated )), format = "pandoc")
}

```



# Mathematical model

## Separate Ordering (SO)
### Not Constained using EOQ

```{r}
n <- length(product_data$`material ID`) #number of materials
cori <- product_data$ordering_cost   #ordering cost for each items
cord <-  1500    #ordering cost whenever there is an order
dei <- product_data$demand_per_year
h<-0.10
box_cost <- product_data$box_cost #pri

# objective function --> I dropped the 1500*62 as it is not decision relevant
obj.fun <- function(q, d= dei, c.or = cori, c.h = h*box_cost  ) (1500*62)+(sum((dei/q)*c.or)+ sum(c.h*(q/2)))
# benchmarks
#obj.fun(product_data$eoq.max)
#obj.fun(product_data$eoq.min)
# constraint function --> also contains the ceiling of lanes and a sum was missing
const.fun <- function(q, bns = product_data$b_not_sorting, bs = product_data$b_sorting, rl = rack_length) {
  sum(bs * ceiling( bns * q / rl))
  }

#const.fun(product_data$eoq.max)
#const.fun(product_data$eoq.min)
# try to figure out a freasible starting solution
#const.fun(product_data$eoq.min/10) -  rack_total_width


obj.const<- data.frame(rbind(c(obj.fun(product_data$eoq.max),const.fun(product_data$eoq.max)),c(obj.fun(product_data$eoq.min),const.fun(product_data$eoq.min) )))

colnames(obj.const)<-c("Cost Function (€)", "Capacity in (mm)") 
rownames(obj.const)<-c("Eoq Max","Eoq Min")

#kable(obj.const,caption = "Cost Values and capacities", format="pandoc")
obj.const
```



### Seperate Ordering With Constraint

using eoq.min or eoq.max in obj.fun yields same results.
```{r}
qopt <- OP(
  objective = F_objective(F=obj.fun ,n=n),
  types = rep("C",n),
  bounds = V_bound(ub= product_data$eoq.min , lb= rep(1, n)),
  constraints = F_constraint(F=const.fun,
                             dir="<=",
                             rhs = rack_total_width)
)

#This shows that minimum EOQ is too big and therefore will not meet the rack space constraint.
#const.fun(min(product_data$eoq.max))#366272
#const.fun(min(product_data$eoq.max)/8)# 61428  still > 56000
#const.fun(min(product_data$eoq.max)/8.7)#52716 < 56000
#const.fun(min(product_data$eoq.max)/10)# 33098 < 56000 is much lower but this can still be the starting value
#round(min(product_data$eoq.max)/10) #15

copt_sol <- ROI_solve(qopt, start = rep(min(product_data$eoq.max)/10,n), solver = "alabama" )
# always check whether the algorithm converged
#copt_sol# The objective value is: 101625.9
# solution
#copt_sol$solution #vector of optimal Quantity that meets the space constraints and minimizes the Obj function.
#round(copt_sol$solution)
#copt_sol$objval #101625.9



#const.fun(copt_sol$solution)#55884
#const.fun(floor(copt_sol$solution))#55092


#obj.fun(copt_sol$solution)#101625.9
#obj.fun(floor(copt_sol$solution))#104639.7  rounded q values

copt_solution<-data.frame(obj.fun(floor(copt_sol$solution)),rbind(floor(copt_sol$solution)))
colnames(copt_solution)<-c("cost",1:62)
rownames(copt_solution)< "constrained"

#kable(copt_solution,caption = "Capacitated values for cost and Quantity",format = "pandoc")

copt_solution
```


### Seperate Ordering no Constraint


```{r}
rack_total_width <- 51200
#qopt1 <- OP(
#  objective = F_objective(F=obj.fun ,n=n),
 # types = rep("C",n),
  #bounds = V_bound(ub= product_data$eoq.max , lb= rep(1, n)),
)

#copt_sol1 <- ROI_solve(qopt1, start = rep(min(product_data$eoq.max)/10,n), solver = "alabama" )
# always check whether the algorithm converged
#copt_sol1$objval
```


```{r}
#copt_noconst<-data.frame(obj.fun(floor(copt_sol1$solution)),rbind(floor(copt_sol1$solution)))
#colnames(copt_noconst)<-c("cost",1:62)
#rownames(copt_noconst)< "unconstrained"
#copt_noconst
```

### EOQ 

unconstrained, just pass in eoq max
```{r}
obj.fun <- function(q, d= dei, c.or = cori, c.h = h*box_cost  ) (1500*62)+(sum((dei/q)*c.or)+ sum(c.h*(q/2)))
obj.fun(product_data$eoq.max)#172279.7
```



```{r}
const.unconst <- data.frame(rbind(c(round(obj.fun(product_data$eoq.max)),round(const.fun(product_data$eoq.max))),c(round( obj.fun(copt_sol$solution)),round(const.fun(copt_sol$solution)))))
colnames(const.unconst)<-c("Cost Function (€)","capacity(mm)")
rownames(const.unconst)<-c("No Capacity","with Capacity")

#kable(const.unconst,caption = "costs",format = "pandoc")
const.unconst
```


# Joint Ordering (JO)



```{r}
n <- length(product_data$`box ID`)
c.or0 <- cord
c.or <- product_data$ordering_cost
#H.vec <- 0.5 * dei * c.sh
# based on the previous definition, I think this should be the vector of holding cost multipliers
#holding cost rate is 10% of price of item, in this case box
H.vec <- 0.1 * product_data$box_cost * dei * 0.5
```


## Objective Function with no capacity constraint

--> an important side note is that the JRP is basically a quite standard EOQ problem which uses the substitution $T_i = \frac{q_i}{y_i}$, i.e., $q_i = T_i \cdot y_i$. Then, you substitute $T_i = m_i \cdot B$ ans result in the JRP (adding the common ordering cost).


```{r}
jrp.obj.fun <- function(m , B, cor, Hvec, cor0) cor0/B + sum(cor/B/m) + sum(Hvec*B*m)

# calculate cycle times
T.vec <- sqrt(c.or/H.vec)
# order products
reo.id <- order(T.vec)
T.vec <- T.vec[reo.id]
c.or <- c.or[reo.id]
H.vec <- H.vec[reo.id]

cost_cycle.mat <- t(cbind(c.or,H.vec,T.vec))
costs.cycle <- data.frame(cost_cycle.mat)
colnames(costs.cycle) <- reo.id 
rownames(costs.cycle) <- c(  "$c_i^{or}$" ,"$H_i$","$T_i$" )
kable(costs.cycle,"pandoc", row.names = T)
#costs.cycle

```




```{r}

# calculate T^2 and cumu. cost shares
res.mat <- t(cbind(c.or/H.vec,(c.or0 + cumsum(c.or))/cumsum(H.vec)))
df <- data.frame(res.mat)
colnames(df) <- reo.id 
rownames(df) <- c(  "$T_i^2$" ,"$T^C$"  )
kable(df,"pandoc", row.names = T)#%>%
  #scroll_box(width = "100%", height = "200px")
```



```{r}
# identify break
which(res.mat[1,] > res.mat[2,])# break occured after 13
id.comb <- min(which(res.mat[1,] > res.mat[2,])) - 1 #3
#id.comb
# calculate B
B <- min(T.vec) #0.01955164
# solution with m - integers #######################
m.vec.int <- round(T.vec/B,0)

m.vec.int[1:id.comb] <- 1  
# re-optimize B for fixed m.vec
B.int <- sqrt(sum(c.or/m.vec.int)/sum(m.vec.int*H.vec))#0.02014868
# total cost 
c.cost.int <- jrp.obj.fun( m = m.vec.int, B=B.int, cor = c.or,Hvec=H.vec, cor0 = c.or0)#192606
df <- data.frame(rbind(round(T.vec/B.int,2), round(T.vec/B.int), m.vec.int))
colnames(df) <- reo.id 

rownames(df) <- c("$m_i= {T_i}/{B}$" ,"$[m_i]$", "$[\\tilde{m}_i]$" )
kable(df,"pandoc", row.names = T)
```


```{r}
q.vec <- m.vec.int*B.int*dei
dei <- dei[reo.id]
df <- data.frame(rbind( m.vec.int, round(m.vec.int*B.int, 2), ceiling(q.vec) ))
colnames(df) <- reo.id 
rownames(df) <- c("$[\\tilde{m}_i]$", "$T_i$", "$q_i$" )
kable(df,"pandoc", row.names = T)
```



```{r}
#q.vec <- m.vec.int*B.int*dei
#paste(round(c.cost.int)," is total cost with no capacity constraint")# 49943
#const.fun(q.vec) > rack_total_width

jrp.no_cap <- data.frame(round(c.cost.int),const.fun(ceiling(q.vec)))

colnames(jrp.no_cap)<-c("Cost(€)","Capacity (mm)")
rownames(jrp.no_cap)<-"Not constrained"

kable(jrp.no_cap,digit=c(1,1),caption = "cost with no capacity constraint", format = "pandoc")
```



## Using T

```{r}
library(ROI)
jrp.obj.fun2 <- function(Tvec, cor, Hvec, cor0) sum(cor0/Tvec) + sum(cor/Tvec) + sum(Hvec*Tvec)
# you need to initialize the parameters in the objective function --> Tvec is to be optimized over, the rest is given
jrp.obj.fun2 <- function(Tvec, cor=c.or, Hvec = H.vec, cor0 = c.or0) sum(cor0/Tvec) + sum(cor/Tvec) + sum(Hvec*Tvec)

qopt2 <- OP(
  objective = F_objective(F=jrp.obj.fun2 , n=62),
  types = rep("C",n),
  bounds = V_bound(ub= rep(30, 62), lb= rep(.001,62))
) 

#jrp.obj.fun2(m.vec.int*B)
#jrp.obj.fun2(rep(.00001,62))
#jrp.obj.fun2(rep(15,62))
# you don't need Alabama here as there is no constraint. Basically you can also optimize this problem by taking derivatives
copt_sol2 <- ROI_solve(qopt2, start = m.vec.int*B)
#copt_sol2

#copt_sol2$solution


jrp.no_cap.cycleTime <- data.frame(round(copt_sol2$objval),const.fun(copt_sol2$solution*dei))

colnames(jrp.no_cap.cycleTime)<-c("Cost(€)","Capacity (mm)")
rownames(jrp.no_cap.cycleTime)<-"Not constrained"

kable(jrp.no_cap.cycleTime,digit=c(1,1),caption = "cost with no capacity constraint using cycle Time", format = "pandoc")

```





## Including capacity constraint


```{r}

# we reinitialize the vectors
n <- length(product_data$`box ID`)
c.or0 <- cord
c.or <- product_data$ordering_cost
dei <- product_data$demand_per_year
H.vec <- 0.1 * product_data$box_cost * dei * 0.5


bnotsort <- product_data$b_not_sorting
bsort <- product_data$b_sorting
copt_sol.solution <- copt_sol$solution



# calculate cycle times
T.vec <- sqrt(c.or/H.vec)
# order products
reo.id <- order(T.vec)
T.vec <- T.vec[reo.id]
c.or <- c.or[reo.id]
H.vec <- H.vec[reo.id]
bnotsort <- bnotsort[reo.id]
bsort <- bsort[reo.id]
copt_sol.solution <- copt_sol.solution[reo.id]
dei <- dei[reo.id]

const.fun2 <- function(m, B = B.start, y= dei, bns = bnotsort, bs = bsort, rl = rack_length) {
  sum(bs * ceiling( bns * m*B*y / rl))
}

T.feas <- copt_sol.solution/dei
B.start <- min(T.feas)

jrp.obj.fun <- function(m , B = B.start, cor = c.or, Hvec = H.vec, cor0 = c.or0) cor0/B + sum(cor/B/m) + sum(Hvec*B*m)
#jrp.obj.fun(m= rep(1, 62))
#const.fun2(m= rep(1, 62))

qopt3 <- OP(
  objective = F_objective(F=jrp.obj.fun ,n=n),
  # now integer decision variables
  types = rep("C",n),
  bounds = V_bound(li = 1:n, ui = 1:n, ub= rep(50, n) , lb= rep(1, n)),
  constraints = F_constraint(F=const.fun2,
                             dir="<=",
                             rhs = rack_total_width)
)
# Good starting point essential --> m = T.feas/B.start
copt_sol3 <- ROI_solve(qopt3, start = T.feas/B.start , solver = "alabama")  # "deoptimr"

#copt_sol3

#copt_sol3$solution

#jrp.obj.fun(m= copt_sol3$solution)
#const.fun2(m= copt_sol3$solution) <= rack_total_width

jrp.cap.m <- data.frame(round(copt_sol3$objval),const.fun2(m= copt_sol3$solution))

colnames(jrp.cap.m)<-c("Cost(€)","Capacity (mm)")
rownames(jrp.cap.m)<-"Capacitated"

kable(jrp.cap.m,digit=c(1,1),caption = "cost with capacity constraint using Basic period approach", format = "pandoc")

```

```{r}
m.start <- floor(copt_sol3$solution) # multiplier m
q.start <- ceiling(m.start * B.start * dei)   # order quantity q (in boxes, rounded up)
#q.start

df2 <- data.frame(rbind( m.start, round(m.start*B.start, 2), ceiling(m.start*B.start*dei) ))
colnames(df2) <- reo.id 
rownames(df2) <- c("$[\\tilde{m}_i]$", "$T_i$", "$q_i$" )
kable(df2,caption = "First 7 items in order before reordering","pandoc", row.names = T)


df2<-df2[mixedorder(colnames(df2))]
kable(df2,caption = "First 7 items in order After reordering","pandoc", row.names = T)
q.start<- as.vector(df2[3,1:62])


#kable(q.start,caption = "After reordering","pandoc", row.names = T)
```





# 1D Cutting Stock Problem

```{r}

```








