---
title: "7. Exercise_Optimization_in_R"
author: "Diego Uchendu"
date: "16/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Exercise 1: 
Formulate the Steiner-Weber model with Euclidean distance and solve it with optim()

$$min_{(u,v)\in R^2} \Rightarrow \sum_{i\in \zeta} (a_i \times \sqrt{(x_i -u)^2+(y_i-v)^2} \space )$$

```{r}
loc.Eucl <- function(loc, a,x,y) sum(a * sqrt((x-loc[1])^2+ (y-loc[2])^2))
n <- 10
a.vec <- sample(1:100, size = n)    # sample weights for each point/customer
x.vec <- rnorm(n)                   # sample x coordinates
y.vec <- rnorm(n)                   # sample y coordinates

res <- optim(par = c(0,0), fn = loc.Eucl, method = "BFGS", a = a.vec, x = x.vec , y= y.vec)
res
```

Substitute optimium location and values in objective function.
```{r}
#using optimum location That is the minimized points
loc= c(0.4130981,0.7379508) #(u,v)
loc.Eucl(loc, a.vec,x.vec,y.vec) # value 463.8937
```

# plotting Steiner Model

```{r}
library(grDevices)        # package for generating color gradients
# generate grid of potential locations
emat <- expand.grid(seq(-3,3, length.out = 100), seq(-3,3, length.out = 100))
# list with grid point coordinates 
plist <- list(x = seq(-3,3, length.out = 100), y = seq(-3,3, length.out = 100))
# evaluate objective function at each potential location
z.vec <- apply(emat, 1, function(x) loc.Eucl(x, a = a.vec, x = x.vec, y = y.vec))
# generate color vector --> assigns a color between green and red depending on the objective value
col.vec <- colorRamp(c("green","orange","red"))( (z.vec-min(z.vec))/(max(z.vec)) )
# recode color vector to RGB code
col.vec <- rgb(col.vec[,1],col.vec[,2],col.vec[,3], maxColorValue = 255)
# create plot
par(mar =c(4,4,0.3,0.3))
plot(emat, xlim=c(-3,3), ylim=c(-3,3), xlab="x", ylab="y", col = col.vec, pch = 15, asp = 1, xaxs = "i", yaxs = "i", cex = 1.5)
points(x.vec, y.vec, pch =20 , cex = log(a.vec)/2)
points(rbind(res$par,res$par), pch =17 , cex = 2, col="blue")
contour(plist, z= matrix(z.vec, ncol=100) , add= TRUE)
legend("topleft", bty="n", legend=c("customer","center"), col=c("black","blue"), pch = c(20,17), bg = "yellow", cex = 1.5)
```

## Exercise 2:

Formulate the Newsvendor model and use optim() to find the optimal order quantity. Compare the optimal solution found by optim() with the theoretical optimum.

The total cost in the normal newsvendor model is given by
$$Z(q)= (c_u + c_o)\times σ \times(φ(q')+q'\times Φ (q'))-c_u \times(q-\mu)$$
with $$q' =\frac{(q-μ)}σ$$ 
Find q while minimizing z
$Z(q)$ =  total cost 
$q$ = order quantity 
$cu$ = Underage cost, cost of losing profit.
$co$ = Overage cost, cost of over ordering.
$\mu$= Mean demand
```{r}
newsven_obj_fun <- function(x, cu,co,mu,sigma){
  q.prime <- (x-mu)/sigma
         (cu + co) * sigma * (dnorm(q.prime) + q.prime * pnorm(q.prime)) - cu * sigma * q.prime

}
#n <- 100
#q.vec <- seq(110, 125, length.out = n) # x axis
#cost.vec <- newsven_obj_fun(x=q.vec, mu = 100, cu=2, co= 0.5, sigma = 20) #yaxis
?optim
min.cost.vec<-optim(par = 1,fn=newsven_obj_fun, "BFGS",mu = 100, cu=2, co= 0.5, sigma = 20 )

min.cost.vec
```

# Exercise 3:

Add an inventory capacity constrint to the Wagner-Whitin model and resolve it with Rglpk.


$$l_t  \le cap \qquad \forall t \in T$$ 
M= sum of demand
```{r}
library(Rglpk)     
# assign names to decision vector
n <- 6                              # number of periods
co <- 50                            # ordering cost rate
ch <- 0.1                           # holding cost rate
d.vec <- round(rnorm(n, mean = 100, sd = 20))   # sample demand
bigM <- sum(d.vec)                  # set big M
x.vec <- numeric(3*n + 1)           # initialize decision variables
# right-hand side vector with 200 units inventory capacity
b.vec <- c(d.vec, rep(0, n + 2), rep(200, n) )  
# objective coefficients 
c.vec <- c(rep(ch, n+1), rep(0, n), rep(co, n))
# vector with constraint directions 
const.vec <- c(rep("==", n), rep("<=", n) , rep("==", 2), rep("<=", n) )
# vector with variable types
vtype.vec <- c(rep("C",n + 1), rep("C",n), rep("B",n)) 
# initialize constraint matrix
A.mat <- matrix(0, ncol = 3*n + 1, nrow = 3*n+2)    
# write coefficient in first n rows (inventory constraints)
for(i in 1:n){
  A.mat[i,i]   <- 1         # coefficient for l_{t-1}
  A.mat[i,i+1] <- -1        # coefficient for l_{t}
  A.mat[i,n + 1 + i] <- 1   # coefficient for q_{t}
}
# write coefficient in rows (n+1):(2*n) (ordering constraints)
for(i in (n+1):(2*n)){
  A.mat[i,i+1] <- 1             # coefficient for q_{t}
  A.mat[i,n + 1 + i] <- -bigM   # coefficient for y_{t}
}
# write coefficient in last two rows (inventory initialization)
A.mat[nrow(A.mat)-n-1,1] <- 1    # coefficient for i_{0}
A.mat[nrow(A.mat)-n,n+1] <- 1    # coefficient for i_{T}
# add inventory constraints
for(i in 1:n){
  A.mat[nrow(A.mat)-n+i, i+1] <- 1
}
# solve MILP
sol <- Rglpk_solve_LP(obj = c.vec, mat = A.mat, dir = const.vec, rhs = b.vec, types = vtype.vec)
# list with solution and demand
list( l = sol$solution[1:(n+1)],         # inventory levels
        q = sol$solution[(n+2):(2*n+1)],   # order quantities
        y = tail(sol$solution, n),         # order indicators
        d = d.vec )                        # demand
# optimum
sol$optimum
```
4. Solve the Transshipment Problem with `Rglpk`:
$$ \min\limits_{x_{ij}} \rightarrow \sum_{i \in \mathcal{V}_a }\sum_{j \in \mathcal{V}_t } c_{ij} \cdot x_{ij} + \sum_{j \in \mathcal{V}_t }\sum_{k \in \mathcal{V}_b } c_{jk} \cdot x_{jk}  $$

$$ \sum_{j \in \mathcal V_t} x_{i,j} \le  a_i \qquad \forall i \in  \mathcal V_a \\ \sum_{j \in \mathcal V_t} x_{j,k} \ge  b_k \qquad  \forall k \in \mathcal V_b  \\ \sum_{i \in \mathcal V_a} x_{ij} = \sum_{k \in \mathcal V_b} x_{x_{jk}} \qquad \forall j \in \mathcal V_t \\ x_{ij} \ge 0$$

where $a_i$ is the maximum supply capacity of node $i \in \mathcal{V}_a$, $b_k$ denotes the demand of node $j \in \mathcal{V}_b$, $c_{ij/jk}$ are transport cost rates and $x_{ij/jk}$ are shipment quantities. 
```{r}
n.a<- 4  #number sources
n.b <- 5 # number sinks
n.t <- 3 #number terminals

nam.A <- paste("A",1:n.a, sep = "") # "A1" "A2" "A3" "A4"
nam.B <- paste("B",1:n.b, sep = "") # "B1" "B2" "B3" "B4" "B5"
nam.T <- paste("T",1:n.t, sep = "") #  "T1" "T2" "T3"
# sample demand & supply

repeat{
  d.vec <- round(runif(n.b, min = 100, max = 200)) # 5 demand values from B nodes
  s.vec <- round(runif(n.a, min = 150, max = 250)) # 4 supply values from A nodes
  #check for sufficient supply
  if(sum(d.vec) < sum(s.vec)) break
}
d.vec # 5 vector values uniformly distributed
s.vec # 4 vector values uniformly distributed
#sample distances
## sources-terminals
d.a.t <- round(runif(n.a*n.t, min = 1, max = 50))
d.a.t # 12 vector values representing quantity from i to j
## terminals-sinks
d.t.b <- round(runif(n.b*n.t, min = 1, max = 30))
d.t.b # 15 vector  values quantitites from j to k(Xjk)
# initialize decision variables
x.vec <- numeric(n.a*n.t + n.b*n.t)
x.vec # 12+15 =27 spaces
# right- hand side vector
?rep()
b.vec <- c(s.vec, d.vec, rep(0, n.t))
b.vec #12 elements in the vector

#objective coefficients
c.vec <- c(d.a.t, d.t.b)
c.vec# 27 elements in the vector

# vector with constraint directions 
const.vec <- c(rep("<=", n.a), rep(">=", n.b) , rep("==", n.t) )
const.vec
# vector with variable types
vtype.vec <- rep("C", length(x.vec))
vtype.vec
# contraint matrix A

```
```{r}
#A Contraint matrix A
A.mat <- matrix(0, ncol = length(x.vec), nrow = n.a + n.b + n.t)
A.mat

# use variable names
colnames(A.mat) <- c(
  apply(expand.grid(nam.A, nam.T), 1, paste, collapse="-"),
  apply(expand.grid(nam.T, nam.B), 1, paste, collapse="-")
)
A.mat # the columns are directions 
```


```{r}
# iteratively add constraints via name matching
# supply constraints


```

#Exercise 5: 
Change the problem by adding constraints that prohibit the center to be closer than 0.5 distance units to any point
```{r}

#solver for non linear, generally contrained programs
library(ROI.plugin.alabama)

n <- 100

#sample weights for each point/customer
a.vec <- sample(1:100, size = n)

#sample x coordinates
x.vec <- rnorm(n)

#sample y coordinates
y.vec <- rnorm(n)

#Set up optimization program
#bounds loc[1]= [-Inf,Inf], loc[2]=[-Inf,Inf] thus the rep(Inf,2) range values for x axis of both loc[1] and loc[2], rep(-Inf,2) range values for y axis of both loc[1] and loc[2]
copt <- OP(
  objective =  F_objective(F=function(loc) sum(a.vec * (abs(x.vec - loc[1]) + abs(y.vec - loc[2]) ) ), n = 2),
  types = rep("C", 2),
  bounds = V_bound(ub = rep(Inf, 2), lb = rep(-Inf, 2)),
  constraints = F_constraint(
    F= function(loc) sum(a.vec * (abs(x.vec - loc[1]) + abs(y.vec - loc[2]) ) ),
    dir = ">=",
    rhs = 0.5)
  
)
#solve the problem
sol.cnst <- ROI_solve(copt, start=c(0,0),solver = "alabama")

#obtain solution
sol <- sol.cnst$solution 
sol # 0.13521671 0.02257355

#objective value
sol.cnst$objval #8717.349
#remember x.vec is a vector this returns the minimum values 
min(abs(x.vec-sol[1])+ abs(y.vec - sol[2]))# 0.214452

#unconstrined solution
constraints(copt) <- NULL
sol.uncnst <- ROI_solve(copt, start=c(0,0))
sol.uncnst# No optimal solution found. The solver message was: No solution. The objective value is: 8.717352e+03

sol.uncnst$objval#8717.352
sol.uncnst$solution#  0.13640996 0.02257363
#compare objective values
sol.cnst$objval - sol.uncnst$objval #-0.002384161

#compare solutions
rbind(sol.cnst$solution, sol.uncnst$solution)

#plot solutions
{plot(x.vec, y.vec, xlab="x", ylab="y", cex = log(a.vec)/4, pch=16, ylim=c(-3,3), xlim=c(-3,3)) 
points(rbind(sol.cnst$solution,sol.cnst$solution), col="blue", cex=2, pch =15)
points(rbind(sol.uncnst$solution,sol.uncnst$solution), col="red", cex=2, pch =17)
legend("topleft", pch=c(15,17), col=c("blue","red"), legend=c("constrained opt.","unconstrained opt."))
}
```
# Exercise 6:
Formulate the Newsvendor model with $\beta$ service level restriction with the `ROI` framework.
```{r}

```
#Exercise 7:

