---
title: "coordinated_ordering4"
author: "GroupB"
date: "24/06/2020"
output:
  html_document: default
  pdf_document: default
---

# Data preparation 

```{r setup, include=TRUE, cache=F}
library(kableExtra)
library("readxl")
product_data <- read_excel("Data_ordering.xlsx",sheet = "product data")
box_data <- read_excel("Data_ordering.xlsx",sheet = "box data")
#rack data
Total_racks=8
levels_per_rack=4
rack_length= 6000
rack_width=1750
rack_height=300

product_data$demand_per_year= ceiling((product_data$`demand per day` *365)/ product_data$`pieces/box`)

product_data$box_cost= product_data$`pieces/box` * product_data$price

product_data

# shortcut ordering cost

product_data <- merge(product_data, box_data[,c("box ID", "ordering cost (€)")], by = "box ID")

colnames(product_data)[8] <- "ordering_cost"

# order_cost <- double(length(product_data$`box ID`))
# 
# for(j in 1: length(box_data$`box ID`)){
#   for (k in 1:length(product_data$`box ID`)) {
#     if(box_data$`box ID`[j]==product_data$`box ID`[k]){
#       order_cost[k] <- box_data$`ordering cost (€)`[j]
#     }
#   }
#   
# }
# 
# product_data$ordering_cost <- order_cost 
# product_data
```





# EOQ modelling


```{r, cache=TRUE}
#dei= demand_per_year, cori= ordering cost for box i
#cord= ordering cost, pri= box cost, h= interest rate

# directly vectorized
so_eoq_fun <- Vectorize(function(dei,cori,cord,pri,h){
  eoq<- sqrt((2*dei*(cori+cord))/(pri*h))
  return(eoq)
})

# maximum EOQ provided there is no coordination of ordering cycles (i.e., cord are not shared among parallely ordered items)
vec_so_eoq_fun.max <- so_eoq_fun(dei = product_data$demand_per_year,cori=product_data$ordering_cost,cord=1500,pri=product_data$box_cost,h=0.10)
# minimum EOQ disregarding common ordering cost cord
vec_so_eoq_fun.min <- so_eoq_fun(dei = product_data$demand_per_year,cori=product_data$ordering_cost,cord=0,pri=product_data$box_cost,h=0.10)

product_data$eoq.min <- round(vec_so_eoq_fun.min)
product_data$eoq.max <- round(vec_so_eoq_fun.max)


```


# Lane occupation 
```{r}
# constraints #####################################
# this can be formulated more elegantly, but it works and that suffices
b_sorting <- double(length(product_data$`box ID`)) 
b_not_sorting <-double(length(product_data$`box ID`))

for(j in 1: length(box_data$`box ID`)){
  for (k in 1:length(product_data$`box ID`)) {
    if(box_data$`box ID`[j]==product_data$`box ID`[k]){
      
      if(box_data$sorting[j]=="width"){
        b_sorting[k] <- box_data$width[j]
        b_not_sorting[k]<- box_data$length[j]
        
      }else{
        b_sorting[k] <- box_data$length[j]
        b_not_sorting[k]<- box_data$width[j]
      }
    }
  }
  
}

product_data$b_sorting <- b_sorting 
product_data$b_not_sorting <- b_not_sorting

lane.min <- ceiling(product_data$eoq.min * product_data$b_not_sorting / rack_length)
lane.max <- ceiling(product_data$eoq.max * product_data$b_not_sorting / rack_length)

# total rack width 

rack_total_width <- rack_width * 4 * 8

# overflow in mm
sum(lane.min*product_data$b_sorting) - rack_total_width
sum(lane.max*product_data$b_sorting) - rack_total_width
# relative overflow in %
(sum(lane.min*product_data$b_sorting) - rack_total_width)/rack_total_width*100
(sum(lane.max*product_data$b_sorting) - rack_total_width)/rack_total_width*100

```

# Constrained EOQ optimization


```{r, cache=TRUE}

library(ROI)
library(ROI.plugin.alabama)

n <- length(product_data$`material ID`) #number of materials
cori <- product_data$ordering_cost   #ordering cost for each items
cord <-  1500    #ordering cost whenever there is an order
dei <- product_data$demand_per_year
h<-0.10
box_cost <- product_data$box_cost #pri

# objective function --> I dropped the 1500*62 as it is not decision relevant
obj.fun <- function(q, d= dei, c.or = cori, c.h = h*box_cost  ) (sum((dei/q)*c.or)+ sum(c.h*(q/2)))
# benchmarks
obj.fun(product_data$eoq.max)
obj.fun(product_data$eoq.min)
# constraint function --> also contains the ceiling of lanes and a sum was missing
const.fun <- function(q, bns = product_data$b_not_sorting, bs = product_data$b_sorting, rl = rack_length) {
  sum(bs * ceiling( bns * q / rl))
  }

const.fun(product_data$eoq.max)
const.fun(product_data$eoq.min)
# try to figure out a freasible starting solution
const.fun(product_data$eoq.min/10) -  rack_total_width

qopt <- OP(
  objective = F_objective(F=obj.fun ,n=n),
  types = rep("C",n),
  bounds = V_bound(ub= product_data$eoq.min , lb= rep(1, n)),
  constraints = F_constraint(F=const.fun,
                             dir="<=",
                             rhs = rack_total_width)
)

#This shows that minimum EOQ is too big and therefore will not meet the rack space constraint.
const.fun(min(product_data$eoq.max))#366272
const.fun(min(product_data$eoq.max)/8)# 61428  still > 56000
const.fun(min(product_data$eoq.max)/8.7)#52716 < 56000
const.fun(round(min(product_data$eoq.max)/10))# 52716 < 56000
round(min(product_data$eoq.max)/10) #17.3


copt_sol <- ROI_solve(qopt, start = rep(min(product_data$eoq.max)/10,n), solver = "alabama" )
# always check whether the algorithm converged
copt_sol# The objective value is: 1.313873e+05
# solution
copt_sol$solution #vector of optimal Quantity that meets the space constraints and minimizes the Obj function.
round(copt_sol$solution)
copt_sol$objval #131387.3

#########################################
# Now you need to fine tune the results 
# --> there are rounding issues -> exactly determining the lane configuration per shelf level
# --> idea to coping with the common ordering cost

const.fun(copt_sol$solution)#55884
const.fun(round(copt_sol$solution))#55686


obj.fun(copt_sol$solution)#131387.3
obj.fun(round(copt_sol$solution))#132096.3  rounded q values

```


# Joint Ordering (JO)

Using Joint replenishment problem

Assumptions:

- One supplier with outbound storage.
- $i=\{1,...,n\}$ products
- Demand rates: $de_i$
- Stock_holding cost rates: $c_i^{sh}$
- Specific setup costs: $c_i^{or}$
- General setup costs: $c^{-or}$
- Cycle time of product $i$: $T_i$

```{r}
n <- length(product_data$`box ID`)
c.or0 <- cord
c.or <- product_data$ordering_cost
#H.vec <- 0.5 * dei * c.sh
# based on the previous definition, I think this should be the vector of holding cost multipliers
H.vec <- 0.1 * product_data$box_cost * dei * .05

```


## Objective Function with no capacity constraint

- $B=$basic cycle time
- $H_i=0.5 \cdot de_i \cdot c_i^{sh}$
- $H_0=0$  Holding cost multiplier are $H_0 \space and  \space H_i$
- $$C(m_i B)=\sum_{i=0}^n (\frac {c_i^{or}}{m_i \cdot B}+H_i \cdot m_i \cdot B) == \frac {c_o^{or}}{B}+ \sum_{i=1}^n (\frac {c_i^{or}}{m_i \cdot B}+H_i \cdot m_i \cdot B)  \\ subject \space to: \quad \quad \quad \quad \quad \quad \quad \quad \quad  \\ m_i \ge m_0 \quad \forall i >0 \\m_i \in \mathbb N \qquad \forall i   $$

--> an important side note is that the JRP is basically a quite standard EOQ problem which uses the substitution $T_i = \frac{q_i}{y_i}$, i.e., $q_i = T_i \cdot y_i$. Then, you substitute $T_i = m_i \cdot B$ ans result in the JRP (adding the common ordering cost).

```{r}
jrp.obj.fun <- function(m , B, cor, Hvec, cor0) cor0/B + sum(cor/B/m) + sum(Hvec*B*m)
```



- *Cycle Time*

we determine $T_i=\sqrt (\frac {c_i^{or}}{H_i})$ and $T^C=\frac {\sum_{j=0}^{i´} c_j^{or}}{\sum_{j=0}^{i´}H_j}$
```{r}
# calculate cycle times
T.vec <- sqrt(c.or/H.vec)
# order products
reo.id <- order(T.vec)
T.vec <- T.vec[reo.id]
c.or <- c.or[reo.id]
H.vec <- H.vec[reo.id]
```

```{r}
# calculate T^2 and cumu. cost shares
res.mat <- t(cbind(c.or/H.vec,(c.or0 + cumsum(c.or))/cumsum(H.vec)))
df <- data.frame(res.mat)
colnames(df) <- reo.id 
rownames(df) <- c(  "$T_i$" ,"$T^C$"  )
kable(df,"pandoc", row.names = T)
```

```{r}
# identify break
which(res.mat[1,] > res.mat[2,])# break occured after 13
id.comb <- min(which(res.mat[1,] > res.mat[2,])) - 1 #3
id.comb
# calculate B
B <- min(T.vec) #0.01955164
# solution with m - integers #######################
m.vec.int <- round(T.vec/B,0)
#I need to address this
m.vec.int[1:id.comb] <- 1  # position 1 to 13 should be assigned 1
# re-optimize B for fixed m.vec
B.int <- sqrt(sum(c.or/m.vec.int)/sum(m.vec.int*H.vec))#0.2431548
# total cost 
c.cost.int <- jrp.obj.fun( m = m.vec.int, B=B.int, cor = c.or,Hvec=H.vec, cor0 = c.or0)#18646.93
df <- data.frame(rbind(round(T.vec/B,2), round(T.vec/B), m.vec.int))
colnames(df) <- reo.id 
rownames(df) <- c("$m_i=\\frac {T_i}{B}$" ,"$[m_i]$", "$[\\tilde{m}_i]$" )
kable(df,"pandoc", row.names = T)
```

Reoptimizing basic cycle time yields $B=$ `r round(B.int,2)` and the total cost are `r round(c.cost.int,2)` 

The order quantities are given by multiplying the cycle times $T_i$ with demand rates $y_i$, i.e. $q_i = T_i \cdot de_i$ such that

```{r}
dei <- dei[reo.id]
df <- data.frame(rbind( m.vec.int, round(m.vec.int*B, 2), round(m.vec.int*B*dei, 2) ))
colnames(df) <- reo.id 
rownames(df) <- c("$[\\tilde{m}_i]$", "$T_i$", "$q_i$" )
kable(df,"pandoc", row.names = T)
```

--> Now the question is: Is this feasible? --> no:

```{r}
q.vec <- m.vec.int*B*dei
const.fun(q.vec) > rack_total_width
```


## Including capacity constraint


When you try to optimize the JRP directly, without introducing the substitution $T_i = m_i \cdot B$ you need to update the JRP function as follows

```{r}
require(ROI)
require(ROI.plugin.alabama)
jrp.obj.fun2 <- function(Tvec, cor, Hvec, cor0) sum(cor0/Tvec) + sum(cor/Tvec) + sum(Hvec*Tvec)
# you need to initialize the parameters in the objective function --> Tvec is to be optimized over, the rest is given
jrp.obj.fun2 <- function(Tvec, cor=c.or, Hvec = H.vec, cor0 = c.or0) sum(cor0/Tvec) + sum(cor/Tvec) + sum(Hvec*Tvec)

qopt2 <- OP(
  objective = F_objective(F=jrp.obj.fun2 , n=62),
  types = rep("C",n),
  bounds = V_bound(ub= rep(30, 62), lb= rep(.001,62))
) 

jrp.obj.fun2(m.vec.int*B)#190854.8
jrp.obj.fun2(rep(.00001,62))#9724500000
jrp.obj.fun2(rep(15,62))#173920.9
# you don't need Alabama here as there is no constraint. Basically you can also optimize this problem by taking derivatives
copt_sol2 <- ROI_solve(qopt2, start = m.vec.int*B)
copt_sol2 #5.671162e+04

copt_sol2$solution

```


--> You should  integrate the shelf capacity constraint 
--> I also recommmend to stick with the basic period approach

This is formally defined as 
$$ \sum_{i=1}^n b_i \cdot \left\lceil\frac{b^{-1}_i \cdot q_i}{rl} \right \rceil \leq \text{tot_rack_width}$$
Thus, with the substitution above, the left-hand side changes to
$$ \sum_{i=1}^n b_i \cdot \left\lceil\frac{b^{-1}_i \cdot T_i \cdot y_i}{rl} \right \rceil = \sum_{i=1}^n b_i \cdot \left\lceil\frac{b^{-1}_i \cdot m_i \cdot B \cdot y_i}{rl} \right \rceil$$

Thus, you can change the constraint function quite easily

```{r}
const.fun2 <- function(m, B = B.start, y= dei, bns = product_data$b_not_sorting, bs = product_data$b_sorting, rl = rack_length) {
  sum(bs * ceiling( bns * m*B*y / rl))
  }
```

Then, you can use the original JRP function and update the model by including the capacity constraint. Therefore I recommend that you fix $B$ to some appropriate value based on the feasible solution above

```{r}
# we reinitialize the vectors
n <- length(product_data$`box ID`)
c.or0 <- cord
c.or <- product_data$ordering_cost
H.vec <- 0.1 * product_data$box_cost * dei * .05


T.feas <- copt_sol$solution/dei
B.start <- min(T.feas)

jrp.obj.fun <- function(m , B = B.start, cor = c.or, Hvec = H.vec, cor0 = c.or0) cor0/B + sum(cor/B/m) + sum(Hvec*B*m)
jrp.obj.fun(m= rep(1, 62))
const.fun2(m= rep(1, 62))

qopt3 <- OP(
  objective = F_objective(F=jrp.obj.fun ,n=n),
  # now integer decision variables
  types = rep("C",n),
  bounds = V_bound(li = 1:n, ui = 1:n, ub= rep(50, n) , lb= rep(1, n)),
  constraints = F_constraint(F=const.fun2,
                             dir="<=",
                             rhs = rack_total_width)
)
# god starting point essential --> m = T.feas/B.start
copt_sol3 <- ROI_solve(qopt3, start = T.feas/B.start , solver = "alabama")

copt_sol3 #2.995570e+05

copt_sol3$solution

jrp.obj.fun(m= copt_sol3$solution) #299557
const.fun2(m= copt_sol3$solution) <= rack_total_width

```

Now you should have some good starting solution. Still you need to round the solution. 
Afterwards, you have to find a layout scheme such that a precise shelf layout results. 


$$\sum_{k=1}^{l=32} \sum_{j=1}^{p_k} lane_{ijk} \cdot b_{ijk} $$

In Each runway, only one part should be stored. Thereby, the width of a runway can be adapted to width of the box type which corresponds to the part assigned to the runway.

, this means total of 62 widths in a pattern and these widths are made up of box id sorting. 

```{r}
df <- data.frame(
  Box_sorting = t(product_data$b_sorting))
df

#some combinations of Box_sorting forms different patterns
```
$rack_{length}=6000 \\$
1. Number of boxes in a lane:
$$n_i=\frac {rl}{b_i^{-1}}=\frac {6000}{297}=20.20=20boxes$$
2. How many lanes part $i$ will occupy if you order some number of boxes
$$lane_i=\frac {q_i}{n_i}\\lane_i=q_i \cdot \frac {b_i^{-1}}{rl}=214 \cdot \frac {297}{6000}=10.593= 11 \space lanes  \\ lane_i=  m_i \cdot B \cdot dei_i \cdot \frac {b_i^{-1}}{rl} $$
$rw= rack \space width$

$\sum_{i=1}^{n=62} b_i=28330$ and $rw=1750$ all in mm, this means that



|$k$\ $i$ $p_k=$Pattern |$b_i=\{b_1, b_2,...,b_{62} \}$|$rw=$ Rack Width|
|---------------|----------------------------|----------|  
|1 | $\{lane_{11},lane_{12},lane_{1,n}\}$ | 1750|
|2 | $\{lane_{21},lane_{22},lane_{2,n}\}$ | 1750|
|3 | $\{lane_{31},lane_{32},lane_{3,n}\}$| 1750|
|k | $\{lane_{k1},lane_{k2},lane_{k,n}\}$| 1750|

1. Sum product of lane for product $i$ in pattern $k$ and box sorting$b_i$ value offer product $i$ in pattern $k$ and $y_{ik} $columns of $b_i$ part of the pattern(0,1) or not.

2. Sum Product of quantity for pattern $k$ for part $i$ and box non sorting for pattern $k$ for part $i$.

$$\sum_{k=1}^{p_k} \sum_{i=1}^{n=62} lane_{ik} \cdot b_{ik} \cdot y_{ik} \le  rw  \qquad  \\ \sum_{i=1}^{n=62} \sum_{k=1}^{p_k} q_{ki} \cdot b_{ki}^{-1} \ge dei_i \\ y \in \{0,1\}$$
The above can be also be reformulated to:

$$ \sum_{i=1}^{n=62} lane_{ik} \cdot b_{ik} \cdot y_{ik} \le  rw  \qquad  \forall k=1,...,p_k\\ \sum_{i=1}^{n=62} \sum_{k=1}^{p_k} q_{ki} \cdot b_{ki}^{-1} \ge dei_i \\ y \in \{0,1\}$$
# Total Formulation:

$$ \min \rightarrow \frac {c_o^{or}}{B}+ \sum_{i=1}^n (\frac {c_i^{or}}{m_i \cdot B}+H_i \cdot m_i \cdot B)$$
Subject to:

$$ \sum_{i=1}^{n=62} m_{ik} \cdot B \cdot dei_{ik} \cdot \frac {b_{ik}^{-1}}{rl} \cdot b_{ik} \cdot y_{ik} \le  rw  \qquad  \forall k=1,...,p_k\\ \sum_{i=1}^{n=62} \sum_{k=1}^{p_k} q_{ki} \cdot b_{ki}^{-1} \ge dei_i \\ y \in \{0,1\}$$
bf
```{r}
sum(product_data$b_sorting)
sum(product_data$b_sorting) > rack_width #true

rack_width
```

```{r}

```

