---
title: "coordinated_ordering4"
author: "GroupB"
date: "24/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DT)
library(kableExtra)
library(knitr)
```

# Data Preprocessing

- *Data Extraction and variable Initialization *:
```{r}
library("readxl")
product_data <- read_excel("Data_ordering.xlsx",sheet = "product data")
box_data <- read_excel("Data_ordering.xlsx",sheet = "box data")

#rack data
Total_racks=8
levels_per_rack=4
rack_length= 6000
rack_width=1750
rack_height=300

product_data$demand_per_year= ceiling((product_data$`demand per day` *365)/ product_data$`pieces/box`)

product_data$box_cost= product_data$`pieces/box` * product_data$price

product_data

product_data <- merge(product_data, box_data[,c("box ID", "ordering cost (€)")], by = "box ID")

colnames(product_data)[8] <- "ordering_cost"

library(dplyr)
head(select(product_data,`box ID`,`material ID`,ordering_cost))
```

- *EOQ*

```{r}
#dei= demand_per_year, cori= ordering cost for box i
#cord= ordering cost, pri= box cost, h= interest rate

# directly vectorized
so_eoq_fun <- Vectorize(function(dei,cori,cord,pri,h){
  eoq<- sqrt((2*dei*(cori+cord))/(pri*h))
  return(eoq)
})

# maximum EOQ provided there is no coordination of ordering cycles (i.e., cord are not shared among parallely ordered items)
vec_so_eoq_fun.max <- so_eoq_fun(dei = product_data$demand_per_year,cori=product_data$ordering_cost,cord=1500,pri=product_data$box_cost,h=0.10)
# minimum EOQ disregarding common ordering cost cord
vec_so_eoq_fun.min <- so_eoq_fun(dei = product_data$demand_per_year,cori=product_data$ordering_cost,cord=0,pri=product_data$box_cost,h=0.10)

product_data$eoq.min <- round(vec_so_eoq_fun.min)
product_data$eoq.max <- round(vec_so_eoq_fun.max)

```


- *Deriving Sorting and non Sorting variables*

```{r}
# constraints #####################################
# this can be formulated more elegantly, but it works and that suffices
b_sorting <- double(length(product_data$`box ID`)) 
b_not_sorting <-double(length(product_data$`box ID`))

for(j in 1: length(box_data$`box ID`)){
  for (k in 1:length(product_data$`box ID`)) {
    if(box_data$`box ID`[j]==product_data$`box ID`[k]){
      
      if(box_data$sorting[j]=="width"){
        b_sorting[k] <- box_data$width[j]
        b_not_sorting[k]<- box_data$length[j]
        
      }else{
        b_sorting[k] <- box_data$length[j]
        b_not_sorting[k]<- box_data$width[j]
      }
    }
  }
  
}

product_data$b_sorting <- b_sorting 
product_data$b_not_sorting <- b_not_sorting
```

- *Lane occupation*:
```{r}


lane.min <- ceiling(product_data$eoq.min * product_data$b_not_sorting / rack_length)
lane.max <- ceiling(product_data$eoq.max * product_data$b_not_sorting / rack_length)

# total rack width 

rack_total_width <- rack_width * 4 * 8

# overflow in mm
sum(lane.min*product_data$b_sorting) - rack_total_width
sum(lane.max*product_data$b_sorting) - rack_total_width
# relative overflow in %
(sum(lane.min*product_data$b_sorting) - rack_total_width)/rack_total_width*100
(sum(lane.max*product_data$b_sorting) - rack_total_width)/rack_total_width*100

```

# Constrained optimization

```{r}
library(ROI)
# solver for non-linear, generally constrained programs
library(ROI.plugin.alabama)

n <- length(product_data$`material ID`) #number of materials
cori <- product_data$ordering_cost   #ordering cost for each items
cord <-  1500    #ordering cost whenever there is an order
dei <- product_data$demand_per_year
h<-0.10
box_cost <- product_data$box_cost #pri

# objective function --> I dropped the 1500*62 as it is not decision relevant
obj.fun <- function(q, d= dei, c.or = cori, c.h = h*box_cost  ) (sum((dei/q)*c.or)+ sum(c.h*(q/2)))
# benchmarks
obj.fun(max(product_data$eoq))
obj.fun(min(product_data$eoq))
# constraint function --> also contains the ceiling of lanes and a sum was missing
const.fun <- function(q, bns = product_data$b_not_sorting, bs = product_data$b_sorting, rl = rack_length) {
  sum(bs * ceiling( bns * q / rl))
  }

const.fun(max(product_data$eoq))
const.fun(min(product_data$eoq))
# try to figure out a freasible starting solution
const.fun(min(product_data$eoq)/10) -  rack_total_width

qopt <- OP(
  objective = F_objective(F=obj.fun ,n=n),
  types = rep("C",n),
  bounds = V_bound(ub=product_data$eoq.max , lb= rep(1, n)),
  constraints = F_constraint(F=const.fun,
                             dir="<=",
                             rhs = rack_total_width)
)

# solve the problem with appropriate starting vlaue and proper solver
?ROI_solve
min(product_data$eoq.max) #173

rack_total_width #56000

#This shows that minimum EOQ is too big and therefore will not meet the rack space constraint.
const.fun(min(product_data$eoq.max))#366272

const.fun(min(product_data$eoq.max)/8)# 61428  still > 56000

const.fun(min(product_data$eoq.max)/8.7)#52716 < 56000

const.fun(round(min(product_data$eoq.max)/10))# 52716 < 56000
round(min(product_data$eoq.max)/10) #17.3

```
We will use min(product_data$eoq)/10 as our starting value because it seemed as the closest value to 56,000 which is the capacity
```{r}
copt_sol <- ROI_solve(qopt, start = rep(min(product_data$eoq.max)/10,n), solver = "alabama" )
# always check whether the algorithm converged
copt_sol# The objective value is: 1.313873e+05
# solution
copt_sol$solution #vector of optimal Quantity that meets the space constraints and minimizes the Obj function.
round(copt_sol$solution)
copt_sol$objval #131387.3

#########################################
# Now you need to fine tune the results 
# --> there are rounding issues -> exactly determining the lane configuration per shelf level
# --> idea to coping with the common ordering cost
```
- *Checks*
```{r}
const.fun(copt_sol$solution)#55884
const.fun(round(copt_sol$solution))#55686

#check obj function

obj.fun(copt_sol$solution)#131387.3
obj.fun(round(copt_sol$solution))#132096.3  rounded q values
```


# Joint Ordering (JO)

Using Joint replenishment problem

Assumptions:
One supplier with outbound storage.
$i=\{1,...,n\}$ products
Demand rates: $de_i$
Stock_holding cost rates: $c_i^{sh}$
Specific setup costs: $c_i^{or}$
General setup costs: $c^{-or}$
Cycle time of product $i$: $T_i$

```{r}
n <- length(product_data$`box ID`)
c.or0 <- cord
c.or <- product_data$ordering_cost
c.sh <- product_data$box_cost
H.vec <- 0.5 * dei * c.sh
```

- *Objective Function with no capacity constraint*

$B=$basic cycle time
$H_i=0.5 * de_i *c_i^{sh}$
$H_0=0$  Holding cost multiplier are $H_0 \space and  \space H_i$
$$C(m_i B)=\sum_{i=0}^n (\frac {c_i^{or}}{m_i \cdot B}+H_i \cdot m_i \cdot B) == \frac {c_o^{or}}{B}+ \sum_{i=1}^n (\frac {c_i^{or}}{m_i \cdot B}+H_i \cdot m_i \cdot B)  \\ subject \space to: \quad \quad \quad \quad \quad \quad \quad \quad \quad  \\ m_i \ge m_0 \quad \forall i >0 \\m_i \in \mathbb N/ \mathbb {2^N} \qquad \forall i   $$



```{r}
jrp.obj.fun <- function(m , B, cor, Hvec, cor0) cor0/B + sum(cor/B/m) + sum(Hvec*B*m)
```

- *Cycle Time*

we determine $T_i=\sqrt (\frac {c_i^{or}}{H_i})$ and $T^C=\frac {\sum_{j=0}^{i´} c_j^{or}}{\sum_{j=0}^{i´}H_j}$
```{r}
# calculate cycle times
T.vec <- sqrt(c.or/H.vec)

# order products
reo.id <- order(T.vec)
T.vec <- T.vec[reo.id]
c.or <- c.or[reo.id]
H.vec <- H.vec[reo.id]
```

```{r}
# calculate T^2 and cumu. cost shares
res.mat <- t(cbind(c.or/H.vec,(c.or0 + cumsum(c.or))/cumsum(H.vec)))

df <- data.frame(res.mat)
colnames(df) <- reo.id 
rownames(df) <- c(  "$T_i$" ,"$T^C$"  )
kable(df,"pandoc", row.names = T)
```

```{r}
# identify break
which(res.mat[1,] > res.mat[2,])# break occured after 13

id.comb <- min(which(res.mat[1,] > res.mat[2,])) - 1 #3
id.comb

# calculate B
B <- min(T.vec) #0.01955164

# solution with m - integers #######################
m.vec.int <- round(T.vec/B,0)

#I need to address this
#m.vec.int[1:id.comb] <- 1  

# re-optimize B for fixed m.vec
B.int <- sqrt(sum(c.or/m.vec.int)/sum(m.vec.int*H.vec))#0.02014868

# total cost 
c.cost.int <- jrp.obj.fun( m = m.vec.int, B=B.int, cor = c.or,Hvec=H.vec, cor0 = c.or0)#192606

df <- data.frame(rbind(round(T.vec/B,2), round(T.vec/B), m.vec.int))
colnames(df) <- reo.id 
rownames(df) <- c("$m_i=\\frac{T_i}{B}$" ,"$[m_i]$", "$[\\tilde{m}_i]$" )
kable(df,"pandoc", row.names = T)
```

Reoptimizing basic cycle time yields $B=$ `r round(B.int,2)` and the total cost are `r round(c.cost.int,2)` 

The order quantities are given by multiplying the cycle times $T_i$ with demand rates $y_i$, i.e. $q_i = T_i \cdot de_i$ such that

```{r}
dei <- dei[reo.id]
df <- data.frame(rbind( m.vec.int, round(m.vec.int*B, 2), round(m.vec.int*B*dei, 2) ))
colnames(df) <- reo.id 
rownames(df) <- c("$[\\tilde{m}_i]$", "$T_i$", "$q_i$" )
kable(df,"pandoc", row.names = T)
```


- *Optimization*:

$$\min \rightarrow \frac {c^{-or}}{T} + \sum_{i=1}^{n=62} (\frac {c_i^{or}}{T}+H_i \cdot T)  $$

- *Optimize before constraint formulation*:

#Error in .check_function_for_sanity(F, n) : cannot evaluate function 'F' using 'n' = 62 parameters

```{r}
jrp.obj.fun2 <- function(T.vec2 , cor,Hvec=H.vec, cor0){
  cor0/T.vec2 + sum(cor/T.vec2) + sum(Hvec*T.vec2)
}

wrapper <- function(x)jrp.obj.fun2(x, cor = c.or,Hvec=H.vec, cor0 = c.or0)



#qopt2 <- OP(
#  objective = F_objective(F=wrapper ,n=62),
#  types = rep("C",n),
#  bounds = V_bound(ub= rep(max(T.vec)), lb= #rep(0,n))
#) 


#Error in .check_function_for_sanity(F, n) : cannot evaluate function 'F' using 'n' = 62 parameters

#copt_sol <- ROI_solve(qopt2, start = rep(min(T.vec),n), solver = "alabama" )

```



```{r}

```

